For Induction-Induction, Induction is Enough

### 2019-11-11 paper/paper.tex
remove infinitary from universaliit.tex

\begin{abstract}
  Inductive-inductive types (IITs) are a generalisation of inductive types in
  type theory. They allow the mutual definition of types with multiple sorts
  where later sorts can be indexed by previous ones. An example is the James
  Chapman-style syntax of type theory with conversion relations for each sort
  where e.g.\ the sort of types is indexed by contexts.  In previous work it was
  shown that all finitary IITs can be constructed from a quotient
  inductive-inductive type (QIIT), namely the theory of IIT signatures. This is
  a small domain-specific type theory where a context is a signature for an
  IIT. In this paper we construct this theory using only inductive types,
  thereby showing a reduction of all finitary IITs to inductive types.  We rely
  on function extensionality and uniqueness of identity proofs. We formalised
  the construction of the theory of IIT signatures in Agda.

@@

@@

We use the naming convention: sorts, operators and equations. Sorts
and operators might have parameters.
\begin{alignat*}{5}
  & \rlap{(1) Substitution calculus} \\[0.5em]
  & \Con && : \Set \\
  & \Ty  && : \Con\ra\Set \\
  & \Sub  && : \Con\ra\Con\ra\Set \\
  & \Tm  && : (\Gamma:\Con)\ra\Ty\,\Gamma\ra\Set \\
  & \id && : \Sub\,\Gamma\,\Gamma \\
  & \blank\circ\blank && : \Sub\,\Theta\,\Delta\ra\Sub\,\Gamma\,\Theta\ra\Sub\,\Gamma\,\Delta \\
  & \ass && : (\sigma \circ \delta) \circ \nu = \sigma \circ (\delta \circ \nu) \\
  & \idl && : \id\circ\sigma = \sigma \\
  & \idr && : \sigma\circ\id = \sigma \\
  & \blank[\blank] && : \Ty\,\Delta\ra\Sub\,\Gamma\,\Delta\ra\Ty\,\Gamma \\
  & \blank[\blank] && : \Tm\,\Delta\,A\ra(\sigma:\Sub\,\Gamma\,\Delta)\ra\Tm\,\Gamma\,(A[\sigma]) \\
  & [\id] && : A[\id] = A \\
  & [\circ] && : A[\sigma \circ \delta] = A[\sigma][\delta] \\
  & [\id] && : t[\id] = t \\
  & [\circ] && : t[\sigma \circ \delta] = t[\sigma][\delta] \\
  & \cdot && : \Con \\
  & \epsilon && : \Sub\,\Gamma\,\cdot \\
  & {\cdot\eta} && : \{\sigma : \Sub\,\Gamma\,\cdot\} \ra \sigma = \epsilon \\
  & \blank\ext\blank && : (\Gamma:\Con)\ra\Ty\,\Gamma\ra\Con \\
  & \blank,\blank && : (\sigma:\Sub\,\Gamma\,\Delta)\ra\Tm\,\Gamma\,(A[\sigma])\ra\Sub\,\Gamma\,(\Delta\ext A) \\
  & \p && : \Sub\,(\Gamma\ext A)\,\Gamma \\
  & \q && : \Tm\,(\Gamma\ext A)\,(A[\p]) \\
  & {\ext\beta_1} && : \p\circ(\sigma, t) = \sigma \\
  & {\ext\beta_2} && : \q[\sigma, t] = t \\
  & {\ext\eta} && : (\p, \q) = \id \\
  & {,\circ} && : (\sigma, t)\circ\delta = (\sigma\circ\delta, t[\delta]) \\[0.5em]
  & \rlap{(2) Sorts and operators} \\[0.5em]
  & \U && : \Ty\,\Gamma \\
  & \El && : \Tm\,\Gamma\,\U\ra\Ty\,\Gamma \\
  & {\U[]} && : \U[\sigma] = \U \\
  & {\El[]} && : (\El\,a)[\sigma] = \El\,(a[\sigma]) \\[0.5em]
  & \rlap{(3) Parameters} \\[0.5em]
  & \Pi && : (a:\Tm\,\Gamma\,\U)\ra\Ty\,(\Gamma\ext\El\,a)\ra\Ty\,\Gamma \\
  & \app && : \Tm\,\Gamma\,(\Pi\,a\,B)\ra\Tm\,(\Gamma\ext \El\,a)\,B \\
  & {\Pi[]} && : (\Pi\,a\,B)[\sigma] = \Pi\,(a[\sigma])\,(B[\sigma^\uparrow]) \\
  & {\app[]} && : (\app\,t)[\sigma^\uparrow] = \app\,(t[\sigma]) \\[0.5em]
  & \rlap{(4) Metatheoretic parameters} \\[0.5em]
  & \Pim && : (T:\Set)\ra(T\ra\Ty\,\Gamma)\ra\Ty\,\Gamma \\
  & \blank\appm \blank && : \Tm\,\Gamma\,(\Pim\,T\,B)\ra(\alpha:T) \ra\Tm\,\Gamma\,(B\,\alpha) \\
  & {\Pim[]} && : (\Pim\,T\,B)[\sigma] = \Pim\,T\,(\lambda \alpha.(B\,\alpha)[\sigma]) \\
  & {\appm[]} && : (t\appm\alpha)[\sigma] = (t[\sigma])\mathop{\appm}\alpha \\[0.5em]
  & \rlap{(5) Infinitary parameters} \\[0.5em]
  & \Pii && : (T:\Set)\ra(T\ra\Tm\,\Gamma\,\U)\ra\Tm\,\Gamma\,\U \\
  & \blank\appi\blank && : \Tm\,\Gamma\,(\El\,(\Pii\,T\,b))\ra(\alpha:T)\ra\Tm\,\Gamma\,(\El\,(b\,\alpha)) \\
  & {\Pii[]} && : (\Pii\,T\,b)[\sigma] = \Pii\,T\,(\lambda \alpha.(b\,\alpha)[\sigma]) \\
  & {\appi[]} && : (t\appi\alpha)[\sigma] = (t[\sigma])\mathop{\appi}\alpha
\end{alignat*}
Abbreviation:
\[
  \sigma^\uparrow = (\sigma\circ\p,\q)
\]

\todo[inline]{add examples where all the different constructors are used}

>>>>

\begin{abstract}
  Inductive-inductive types (IITs) are a generalisation of inductive types in
  type theory. They allow the mutual definition of types with multiple sorts
  where later sorts can be indexed by previous ones. An example is the James
  Chapman-style syntax of type theory with conversion relations for each sort
  where e.g.\ the sort of types is indexed by contexts. It follows from previous
  work that all finitary IITs can be constructed from a quotient
  inductive-inductive type (QIIT), namely the theory of IIT signatures. This is
  a small domain-specific type theory where a context is a signature for an
  IIT. In this paper we construct this theory using only inductive types,
  thereby showing a reduction of all finitary IITs to inductive types.  We rely
  on function extensionality and uniqueness of identity proofs. We formalised
  the construction of the theory of IIT signatures in Agda.

@@

\todo{András: overall todos: introduction, section 2 examples, intro to section 3, future work, maybe changing big Sigma notation to (x : A) x B notation (or changing the other way?)n}

@@

We use the naming convention: sorts, operators and equations. Sorts
and operators might have parameters.
\begin{alignat*}{5}
  & \rlap{(1) Substitution calculus} \\[0.5em]
  & \Con && : \Set \\
  & \Ty  && : \Con\ra\Set \\
  & \Sub  && : \Con\ra\Con\ra\Set \\
  & \Tm  && : (\Gamma:\Con)\ra\Ty\,\Gamma\ra\Set \\
  & \id && : \Sub\,\Gamma\,\Gamma \\
  & \blank\circ\blank && : \Sub\,\Theta\,\Delta\ra\Sub\,\Gamma\,\Theta\ra\Sub\,\Gamma\,\Delta \\
  & \ass && : (\sigma \circ \delta) \circ \nu = \sigma \circ (\delta \circ \nu) \\
  & \idl && : \id\circ\sigma = \sigma \\
  & \idr && : \sigma\circ\id = \sigma \\
  & \blank[\blank] && : \Ty\,\Delta\ra\Sub\,\Gamma\,\Delta\ra\Ty\,\Gamma \\
  & \blank[\blank] && : \Tm\,\Delta\,A\ra(\sigma:\Sub\,\Gamma\,\Delta)\ra\Tm\,\Gamma\,(A[\sigma]) \\
  & [\id] && : A[\id] = A \\
  & [\circ] && : A[\sigma \circ \delta] = A[\sigma][\delta] \\
  & [\id] && : t[\id] = t \\
  & [\circ] && : t[\sigma \circ \delta] = t[\sigma][\delta] \\
  & \cdot && : \Con \\
  & \epsilon && : \Sub\,\Gamma\,\cdot \\
  & {\cdot\eta} && : \{\sigma : \Sub\,\Gamma\,\cdot\} \ra \sigma = \epsilon \\
  & \blank\ext\blank && : (\Gamma:\Con)\ra\Ty\,\Gamma\ra\Con \\
  & \blank,\blank && : (\sigma:\Sub\,\Gamma\,\Delta)\ra\Tm\,\Gamma\,(A[\sigma])\ra\Sub\,\Gamma\,(\Delta\ext A) \\
  & \p && : \Sub\,(\Gamma\ext A)\,\Gamma \\
  & \q && : \Tm\,(\Gamma\ext A)\,(A[\p]) \\
  & {\ext\beta_1} && : \p\circ(\sigma, t) = \sigma \\
  & {\ext\beta_2} && : \q[\sigma, t] = t \\
  & {\ext\eta} && : (\p, \q) = \id \\
  & {,\circ} && : (\sigma, t)\circ\delta = (\sigma\circ\delta, t[\delta]) \\[0.5em]
  & \rlap{(2) Sorts and operators} \\[0.5em]
  & \U && : \Ty\,\Gamma \\
  & \El && : \Tm\,\Gamma\,\U\ra\Ty\,\Gamma \\
  & {\U[]} && : \U[\sigma] = \U \\
  & {\El[]} && : (\El\,a)[\sigma] = \El\,(a[\sigma]) \\[0.5em]
  & \rlap{(3) Parameters} \\[0.5em]
  & \Pi && : (a:\Tm\,\Gamma\,\U)\ra\Ty\,(\Gamma\ext\El\,a)\ra\Ty\,\Gamma \\
  & \app && : \Tm\,\Gamma\,(\Pi\,a\,B)\ra\Tm\,(\Gamma\ext \El\,a)\,B \\
  & {\Pi[]} && : (\Pi\,a\,B)[\sigma] = \Pi\,(a[\sigma])\,(B[\sigma^\uparrow]) \\
  & {\app[]} && : (\app\,t)[\sigma^\uparrow] = \app\,(t[\sigma]) \\[0.5em]
  & \rlap{(4) Metatheoretic parameters} \\[0.5em]
  & \Pim && : (T:\Set)\ra(T\ra\Ty\,\Gamma)\ra\Ty\,\Gamma \\
  & \blank\appm \blank && : \Tm\,\Gamma\,(\Pim\,T\,B)\ra(\alpha:T) \ra\Tm\,\Gamma\,(B\,\alpha) \\
  & {\Pim[]} && : (\Pim\,T\,B)[\sigma] = \Pim\,T\,(\lambda \alpha.(B\,\alpha)[\sigma]) \\
  & {\appm[]} && : (t\appm\alpha)[\sigma] = (t[\sigma])\mathop{\appm}\alpha \\[0.5em]
\end{alignat*}
\todo{We may need more abbrevs here}
Abbreviation:
\[
  \sigma^\uparrow = (\sigma\circ\p,\q)
\]

\todo{add examples where all the different constructors are used}
